### 2015-06-30 Appendix/Appendix1.Rnw
Major changes following committee discussion in defense.

% Appendix1 file from standard thesis template
\appendixtitle
\appendix
\chapter{Appendix for Chapter 2}

@@

The following appendix contains the questions and point rubrics pair of ARTIST scaled questions followed by the pair of applied theory-based inference problems as they appeared on the final exam.

@@

For the ARTIST Model described in Subsection~\ref{ArtistModel} and the Applied Model described in Subsection~\ref{AppliedModel} a set of model diagnostics were conducted. Each of these MANCOVA models are parameterized as specified in Section~\ref{analysis}; as such the assessment of model assumptions will be very similar for each model. First, repercussions of violating the assumption of independence are assessed through a simulation study.  Residual plots will be used to assess the assumptions of linearity and constant variance.  The univariate and bivariate normality of the error terms will be assessed visually using normal quantile plots and a scatterplot of the paired residuals from the model. Additionally, although it is not a modeling assumption, MANCOVA models are best behaved with low to moderate correlation between the response variables, because then then model is able to capture variance unique to each response.

\subsection{Simulation Study for Violations of Independence}
\label{simappend}

The assumption of independence of errors between students in the bivariate MANCOVA models specified in  (\ref{eq:mancova}), which we employed in Sections~\ref{ArtistModel}~and~\ref{AppliedModel} , is made in order to fit models with parsimonious variance structures. The assumption of independence between students is very likely violated in practice, because learning outcomes for students attending the same class with the same instructors in the same semester could well be related. It is therefore important to assess the consequences this violation has on the results of the MANCOVA models fit in the paper.  We wish to identify the effect of the violation on the error rates in the tests for curriculum effects on the learning outcomes. We elected to explore the consequences through a simulation study wherein the assumption of independence was knowingly violated and the effects on errors rates were recorded.

We chose the ARTIST Model (\ref{eq:mancova}) from Sections~\ref{ArtistModel} as the basis for our simulation study. Under the assumption of independence in the errors between students, we found weak evidence of a curriculum effect on the bivariate learning outcomes based on Pillai's $\lambda$. Further inspection of the individual responses using t-tests revealed evidence of a significant effect of the curriculum on confidence interval related outcomes, but no evidence of a curriculum effect on hypothesis test related outcomes.  In order to assess the reliability of the results, we must first know how a violation of the assumption of independence would impact the error rates for these tests.  We simulate under the scenario where the curriculum effect on learning outcomes is non-existent and assess the effect on   error rates under differing degrees of violation.

The ARTIST Model of the form  (\ref{eq:mancova}) assumes  that errors from different students are independent, i.e.~$\Cov[\epsilon_{ik}, \epsilon_{jl}] = 0$, for all $i \neq j$ with $ 1 \le i,j \le n$. In our simulation setting, we simulate data from a model adapted from the model~(\ref{eq:mancova}), where there is no curriculum effect, i.e.~$\tau_{1} =  \tau_{2} = 0$ and where additionally the assumption of independent errors is violated by assuming a between student covariance of 
\[
\Cov[\epsilon_{ik}, \epsilon_{jl}] = \eta \sigma_{kl}^2, \ \ \ \text{ where } \eta \in [0,1]  \text{ for all } i \ne j. 
\]
$\eta$ scales the covariance between all students uniformly and can be adjusted to increase the violation of independence. Note that when $\eta = 0$ we have no violation of independence. Thus our generative model for the ARTIST responses are linear combinations of the midterm and lab~5 scores with correlated errors. This generative model for simulation can be organized using vector notation as follows:

@@

\begin{eqnarray}\label{eq:linear}
\begin{bmatrix}
  \vec{y}_1 \\ \vec{y}_2
  \end{bmatrix}  = 
% %---- 
\begin{bmatrix}
 y_{11} \\ \vdots \\ y_{n1} \\ y_{12} \\ \vdots \\ y_{n2} 
\end{bmatrix}  =
% %----  
 \begin{bmatrix}
  \beta_{01} + x_{11}\beta_{11} + x_{12}\beta_{21}  \\
  \vdots  \\ 
  \beta_{01} + x_{n1}\beta_{11} + x_{n2}\beta_{21}  \\
  \beta_{02} + x_{11}\beta_{12} + x_{12}\beta_{22}  \\
  \vdots  \\ 
  \beta_{02} + x_{n1}\beta_{12} + x_{n2}\beta_{22}  \\
 \end{bmatrix} +
 \begin{bmatrix}
 \epsilon_{11} \\ \vdots \\ \epsilon_{n1} \\ \epsilon_{12} \\ \vdots \\ \epsilon_{n2} 
 \end{bmatrix} = 
 % %----
 \begin{bmatrix}
  X\vec{\beta}_1 \\ X\vec{\beta}_2
  \end{bmatrix} + 
  \vec{\epsilon},
\end{eqnarray} 

@@

\begin{eqnarray}\label{eq:error}
\vec{\epsilon} \sim
% %----  
 \text{MVN}\left(
\vec{0} \hspace{.1cm}
  , 
  \begin{bmatrix}
   \sigma_{11}^2 & \sigma_{12}^2 \\ 
   \sigma_{21}^2 & \sigma_{22}^2
  \end{bmatrix}
\otimes
 \left( \eta 1_{n \times n} + (1- \eta) I_{n \times n}\right)
%  \left[\begin{array}{ccccc|ccccc}
%   \sigma_{11}^2 & \eta \sigma_{11}^2 & \cdots & \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2\\
%   \eta \sigma_{11}^2 & \sigma_{11}^2 & \ddots & \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \eta \sigma_{12}^2 & \sigma_{12}^2 & \ddots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2\\
% \vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
% \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \ddots & \sigma_{11}^2 & \eta \sigma_{11}^2 & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \ddots & \sigma_{12}^2 & \eta \sigma_{12}^2\\
% \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \cdots & \eta \sigma_{11}^2 & \sigma_{11}^2 & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \sigma_{12}^2\\

@@

% \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \sigma_{22}^2 & \eta \sigma_{22}^2 & \cdots & \eta \sigma_{22}^2 & \eta \sigma_{22}^2\\
%   \eta \sigma_{12}^2 & \sigma_{12}^2 & \ddots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \eta \sigma_{22}^2 & \sigma_{22}^2 & \ddots & \eta \sigma_{22}^2 & \eta \sigma_{22}^2\\
% \vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
% \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \ddots & \sigma_{12}^2 & \eta \sigma_{12}^2 & \eta \sigma_{22}^2 & \eta \sigma_{22}^2 & \ddots & \sigma_{22}^2 & \eta \sigma_{22}^2\\
% \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \sigma_{12}^2 & \eta \sigma_{22}^2 & \eta \sigma_{22}^2 & \cdots & \eta \sigma_{22}^2 & \sigma_{22}^2\\  
%  \end{array}\right]
 \right),
\end{eqnarray}
where $A \otimes B = \left( a_{ij} B \right)$ is the Kronecker product of matrices $A$ and $B$. $\vec{0}$ is the zero vector of length $2n$, $1_{n\times n}$ is the $n \times n$ matrix consisting of 1s only, and $I_{n \times n}$ is the identity matrix of dimension $n \times n$.
Here, $x_{i1}$ is the midterm exam score for student $i$ and $x_{i2}$ is the lab~5 score for student $i$. In order to achieve practically relevant values for the coefficients $\beta_{kp}$ of the generative model we selected the estimated values from a MANCOVA model fit to the ARTIST responses using only the midterm exam and lab~5 scores.  Likewise, the variance components, $\sigma_{kl}$ for the generative model were selected using the estimated variance structure from the same reduced MANCOVA model for the ARTIST responses.

The simulation procedure for assessing the error rates under violation independence between student responses was conducted in a five step process.
For each $\eta$ in $\{0,0.1,0.2,0.3,0.4,0.5\}$ repeat the following process for $M = 25,000$ simulations each:
%
\begin{enumerate}
\item Simulate an error vector, $\vec{\epsilon}^{\hspace{.1cm}(m)}$, from the multivariate normal distribution as defined in~(\ref{eq:error}).
\item Derive corresponding responses based on model (\ref{eq:linear}) as:
%
\[
\begin{bmatrix}
  \vec{y}_1^{\hspace{.1cm}(m)} \\ \vec{y}_2^{\hspace{.1cm}(m)}
  \end{bmatrix}   =
   % %----
 \begin{bmatrix}
  X\vec{\beta}_1 \\ X\vec{\beta}_2
  \end{bmatrix} + 
  \vec{\epsilon}^{\hspace{.1cm}(m)}
\]
%
\item Create assignment to curricula  for each student randomly.

\item Fit the ARTIST Model (assuming independence) as given in model (\ref{eq:mancova}) to the simulated responses
\item Record $\hat{\beta}_{kp}^{\hspace{.1cm}(m)}$, $\text{SE}\left[\hat{\beta}_{kp}^{\hspace{.1cm}(m)}\right]$, $\tau_k^{\hspace{.1cm}(m)}$, $\text{SE}\left[\tau_k^{\hspace{.1cm}(m)}\right]$ and $\hat{\sigma}_{kl}^{\hspace{.1cm}(m)}$ $\forall k,l \in \{1,2\}$ and $p \in \{0,1,2\}$. \\
Also record Pillai's $\lambda^{\hspace{.1cm}(m)}$ testing for overall curriculum effect.
%\item Repeat simulation steps 1 through 6, $M=25000$ times

\end{enumerate}

<<results, echo=FALSE, include=FALSE>>=
load("SimulationResults.RData")
library(plyr)
eta <- (0:5)/10
results <- ldply(1:6, function(x) data.frame(eta=eta[x], resultlist[[x]]))
# beta1s and beta2s are actually tau1 and tau2
simresults = ddply(results, .(eta), summarise,
      trt1=mean(beta1s),
      trt2=mean(beta2s),
      type1.1=mean(pval1s<0.05),
      type1.2=mean(pval2s<0.05),      
      betamidterm1s = mean(betamidterm1s),
      pvalmidterms1s = mean(pvalmidterm1s),
      sebetamidterm1s = mean(SEbetamidterm1s),
      betamidterm2s = mean(betamidterm2s),
      pvalmidterms2s = mean(pvalmidterm2s),
      sebetamidterm2s = mean(SEbetamidterm2s),
      avgpillai = mean(pillais),
      avgpillaipval = mean(pillaipvals),
      sebeta1s = mean(SEbeta1s),
      sebeta2s = mean(SEbeta1s),
      sdbeta1 = sd(beta1s),
      sdbeta2 = sd(beta2s)
)

L <- length(resultlist)
M = nrow(resultlist[[1]]) 
errorsummary <- data.frame(T2ErrorRateBetaMidterm1 = rep(NA, L),
                           T2ErrorRateBetaMidterm2 = rep(NA, L),
                           T2ErrorRateBetaLab1 = rep(NA, L),
                           T2ErrorRateBetaLab2 = rep(NA, L),
                           T1ErrorRateBeta1 = rep(NA, L),
                           T1ErrorRateBeta2 = rep(NA, L),
                           T1ErrorRatePillai = rep(NA, L))

for (l in 1:L){
  errorsummary[l,1] <- 1- length(which(resultlist[[l]]$pvalmidterm1s <= 0.05)) / M
  errorsummary[l,2] <- 1- length(which(resultlist[[l]]$pvalmidterm2s <= 0.05)) / M
  errorsummary[l,3] <- 1- length(which(resultlist[[l]]$pvallab1s <= 0.05)) / M
  errorsummary[l,4] <- 1- length(which(resultlist[[l]]$pvallab2s <= 0.05)) / M
  errorsummary[l,5] <- length(which(resultlist[[l]]$pval1s <= 0.05)) / M
  errorsummary[l,6] <- length(which(resultlist[[l]]$pval2s <= 0.05)) / M
  errorsummary[l,7] <- length(which(resultlist[[l]]$pillaipvals <= 0.05)) / M
}
errorsummary$eta <- eta
#errorsummary
@

Histograms for the estimated parameters from simulations at each value of $\eta$ are displayed in Figure~\ref{fig:simhistogrid}.  Note that the histograms are each centered around the true values of the parameters from the generative model, indicating that the estimates remain unbiased as the violation of independence between student responses increases. The figure also displays a decreasing variability in the parameter estimates as $\eta$ increases. The standard deviation for the treatment effect estimates, $s_{\tau_k}$, can be found in table~\ref{tab:simvarcompare}; along with averages from standard errors for the treatment effects, $\hat{\text{SE}}\left[\tau_k\right]$.  Note that while the former is measuring the variability of the coefficients from all simulations, the latter is the average of variability estimates from each individual simulation.  The paired values are nearly equivalent, meaning that the standard errors used in the t-tests for treatment effects shrink at the same rate as the true sampling variability in the parameters estimates as the violation of independence increases.

>>>>

% Appendix1 file from standard thesis template
\chapter{ }

@@

The following appendix contains the questions and point rubrics for the ARTIST scaled questions and the applied theory-based inference problems as they appeared on the final exam.

@@

For the ARTIST Model described in Subsection~\ref{ArtistModel} and the Applied Model described in Subsection~\ref{AppliedModel} a set of model diagnostics were conducted. Each of these MANCOVA models are parameterized as specified in Section~\ref{analysis}; as such the assessment of model assumptions will be very similar for each model.   Residual plots will be used to assess the assumptions of linearity and constant variance.  The univariate and bivariate normality of the error terms will be assessed visually using normal quantile plots and a scatterplot of the paired residuals from the model. Additionally, although it is not a modeling assumption, MANCOVA models are best behaved with low to moderate correlation between the response variables, because then then model is able to capture variance unique to each response.

% \subsection{Simulation Study for Violations of Independence}
% \label{simappend}

@@

% \km{To be removed after new simulation section added to main body.}\\

@@

% The ARTIST Model of the form (\ref{eq:mancova}) assumes  that errors from different students are independent, i.e.~$\Cov[\epsilon_{ik}, \epsilon_{jl}] = 0$, for all $i \neq j$ with $ 1 \le i,j \le n$. In our simulation setting, we simulate data from a model adapted from the model~(\ref{eq:mancova}), where there is no curriculum effect, i.e.~$\tau_{1} =  \tau_{2} = 0$ and where additionally the assumption of independent errors is violated by assuming a between student covariance of 
% \[
% \Cov[\epsilon_{ik}, \epsilon_{jl}] = \eta \sigma_{kl}^2, \ \ \ \text{ where } \eta \in [0,1]  \text{ for all } i \ne j. 
% \]
% $\eta$ scales the covariance between all students uniformly and can be adjusted to increase the violation of independence. Note that when $\eta = 0$ we have no violation of independence. Thus our generative model for the ARTIST responses are linear combinations of the midterm and lab~5 scores with correlated errors. This generative model for simulation can be organized using vector notation as follows:
% %
% \begin{eqnarray}\label{eq:linear}
% \begin{bmatrix}
%   \vec{y}_1 \\ \vec{y}_2
%   \end{bmatrix}  = 
% % %---- 
% \begin{bmatrix}
%  y_{11} \\ \vdots \\ y_{n1} \\ y_{12} \\ \vdots \\ y_{n2} 
% \end{bmatrix}  =
% % %----  
%  \begin{bmatrix}
%   \beta_{01} + x_{11}\beta_{11} + x_{12}\beta_{21}  \\
%   \vdots  \\ 
%   \beta_{01} + x_{n1}\beta_{11} + x_{n2}\beta_{21}  \\
%   \beta_{02} + x_{11}\beta_{12} + x_{12}\beta_{22}  \\
%   \vdots  \\ 
%   \beta_{02} + x_{n1}\beta_{12} + x_{n2}\beta_{22}  \\
%  \end{bmatrix} +
%  \begin{bmatrix}
%  \epsilon_{11} \\ \vdots \\ \epsilon_{n1} \\ \epsilon_{12} \\ \vdots \\ \epsilon_{n2} 
%  \end{bmatrix} = 
%  % %----
%  \begin{bmatrix}
%   X\vec{\beta}_1 \\ X\vec{\beta}_2
%   \end{bmatrix} + 
%   \vec{\epsilon},
% \end{eqnarray} 
% %
% \begin{eqnarray}\label{eq:error}
% \vec{\epsilon} \sim
% % %----  
%  \text{MVN}\left(
% \vec{0} \hspace{.1cm}
%   , 
%   \begin{bmatrix}
%    \sigma_{11}^2 & \sigma_{12}^2 \\ 
%    \sigma_{21}^2 & \sigma_{22}^2
%   \end{bmatrix}
% \otimes
%  \left( \eta 1_{n \times n} + (1- \eta) I_{n \times n}\right)
% %  \left[\begin{array}{ccccc|ccccc}
% %   \sigma_{11}^2 & \eta \sigma_{11}^2 & \cdots & \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2\\
% %   \eta \sigma_{11}^2 & \sigma_{11}^2 & \ddots & \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \eta \sigma_{12}^2 & \sigma_{12}^2 & \ddots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2\\
% % \vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
% % \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \ddots & \sigma_{11}^2 & \eta \sigma_{11}^2 & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \ddots & \sigma_{12}^2 & \eta \sigma_{12}^2\\
% % \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \cdots & \eta \sigma_{11}^2 & \sigma_{11}^2 & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \sigma_{12}^2\\
% % \hline
% % \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \sigma_{22}^2 & \eta \sigma_{22}^2 & \cdots & \eta \sigma_{22}^2 & \eta \sigma_{22}^2\\
% %   \eta \sigma_{12}^2 & \sigma_{12}^2 & \ddots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \eta \sigma_{22}^2 & \sigma_{22}^2 & \ddots & \eta \sigma_{22}^2 & \eta \sigma_{22}^2\\
% % \vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
% % \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \ddots & \sigma_{12}^2 & \eta \sigma_{12}^2 & \eta \sigma_{22}^2 & \eta \sigma_{22}^2 & \ddots & \sigma_{22}^2 & \eta \sigma_{22}^2\\
% % \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \sigma_{12}^2 & \eta \sigma_{22}^2 & \eta \sigma_{22}^2 & \cdots & \eta \sigma_{22}^2 & \sigma_{22}^2\\  
% %  \end{array}\right]
%  \right),
% \end{eqnarray}
% where $A \otimes B = \left( a_{ij} B \right)$ is the Kronecker product of matrices $A$ and $B$. $\vec{0}$ is the zero vector of length $2n$, $1_{n\times n}$ is the $n \times n$ matrix consisting of 1s only, and $I_{n \times n}$ is the identity matrix of dimension $n \times n$.
% Here, $x_{i1}$ is the midterm exam score for student $i$ and $x_{i2}$ is the lab~5 score for student $i$. In order to achieve practically relevant values for the coefficients $\beta_{kp}$ of the generative model we selected the estimated values from a MANCOVA model fit to the ARTIST responses using only the midterm exam and lab~5 scores.  Likewise, the variance components, $\sigma_{kl}$ for the generative model were selected using the estimated variance structure from the same reduced MANCOVA model for the ARTIST responses. 
% The simulation procedure for assessing the error rates under violation independence between student responses was conducted in a five step process.
% For each $\eta$ in $\{0,0.1,0.2,0.3,0.4,0.5\}$ repeat the following process for $M = 25,000$ simulations each:
% %
% \begin{enumerate}
% \item Simulate an error vector, $\vec{\epsilon}^{\hspace{.1cm}(m)}$, from the multivariate normal distribution as defined in~(\ref{eq:error}).
% \item Derive corresponding responses based on model (\ref{eq:linear}) as:
% %
% \[
% \begin{bmatrix}
%   \vec{y}_1^{\hspace{.1cm}(m)} \\ \vec{y}_2^{\hspace{.1cm}(m)}
%   \end{bmatrix}   =
%    % %----
%  \begin{bmatrix}
%   X\vec{\beta}_1 \\ X\vec{\beta}_2
%   \end{bmatrix} + 
%   \vec{\epsilon}^{\hspace{.1cm}(m)}
% \]
% %
% \item Create assignment to curricula  for each student randomly. 
% \item Fit the ARTIST Model (assuming independence) as given in model (\ref{eq:mancova}) to the simulated responses
% \item Record $\hat{\beta}_{kp}^{\hspace{.1cm}(m)}$, $\text{SE}\left[\hat{\beta}_{kp}^{\hspace{.1cm}(m)}\right]$, $\tau_k^{\hspace{.1cm}(m)}$, $\text{SE}\left[\tau_k^{\hspace{.1cm}(m)}\right]$ and $\hat{\sigma}_{kl}^{\hspace{.1cm}(m)}$ $\forall k,l \in \{1,2\}$ and $p \in \{0,1,2\}$. \\
% Also record Pillai's $\lambda^{\hspace{.1cm}(m)}$ testing for overall curriculum effect.
% %\item Repeat simulation steps 1 through 6, $M=25000$ times
% \end{enumerate}
% <<results, echo=FALSE, include=FALSE>>=
% load("SimulationResults.RData")
% library(plyr)
% eta <- (0:5)/10
% results <- ldply(1:6, function(x) data.frame(eta=eta[x], resultlist[[x]]))
% # beta1s and beta2s are actually tau1 and tau2
% simresults = ddply(results, .(eta), summarise,
%       trt1=mean(beta1s),
%       trt2=mean(beta2s),
%       type1.1=mean(pval1s<0.05),
%       type1.2=mean(pval2s<0.05),      
%       betamidterm1s = mean(betamidterm1s),
%       pvalmidterms1s = mean(pvalmidterm1s),
%       sebetamidterm1s = mean(SEbetamidterm1s),
%       betamidterm2s = mean(betamidterm2s),
%       pvalmidterms2s = mean(pvalmidterm2s),
%       sebetamidterm2s = mean(SEbetamidterm2s),
%       avgpillai = mean(pillais),
%       avgpillaipval = mean(pillaipvals),
%       sebeta1s = mean(SEbeta1s),
%       sebeta2s = mean(SEbeta1s),
%       sdbeta1 = sd(beta1s),
%       sdbeta2 = sd(beta2s)
% )
% L <- length(resultlist)
% M = nrow(resultlist[[1]]) 
% errorsummary <- data.frame(T2ErrorRateBetaMidterm1 = rep(NA, L),
%                            T2ErrorRateBetaMidterm2 = rep(NA, L),
%                            T2ErrorRateBetaLab1 = rep(NA, L),
%                            T2ErrorRateBetaLab2 = rep(NA, L),
%                            T1ErrorRateBeta1 = rep(NA, L),
%                            T1ErrorRateBeta2 = rep(NA, L),
%                            T1ErrorRatePillai = rep(NA, L))
% for (l in 1:L){
%   errorsummary[l,1] <- 1- length(which(resultlist[[l]]$pvalmidterm1s <= 0.05)) / M
%   errorsummary[l,2] <- 1- length(which(resultlist[[l]]$pvalmidterm2s <= 0.05)) / M
%   errorsummary[l,3] <- 1- length(which(resultlist[[l]]$pvallab1s <= 0.05)) / M
%   errorsummary[l,4] <- 1- length(which(resultlist[[l]]$pvallab2s <= 0.05)) / M
%   errorsummary[l,5] <- length(which(resultlist[[l]]$pval1s <= 0.05)) / M
%   errorsummary[l,6] <- length(which(resultlist[[l]]$pval2s <= 0.05)) / M
%   errorsummary[l,7] <- length(which(resultlist[[l]]$pillaipvals <= 0.05)) / M
% }
% errorsummary$eta <- eta
% #errorsummary
% @
% Histograms for the estimated parameters from simulations at each value of $\eta$ are displayed in Figure~\ref{fig:simhistogrid}.  Note that the histograms are each centered around the true values of the parameters from the generative model, indicating that the estimates remain unbiased as the violation of independence between student responses increases. The figure also displays a decreasing variability in the parameter estimates as $\eta$ increases. The standard deviation for the treatment effect estimates, $s_{\tau_k}$, can be found in table~\ref{tab:simvarcompare}; along with averages from standard errors for the treatment effects, $\hat{\text{SE}}\left[\tau_k\right]$.  Note that while the former is measuring the variability of the coefficients from all simulations, the latter is the average of variability estimates from each individual simulation.  The paired values are nearly equivalent, meaning that the standard errors used in the t-tests for treatment effects shrink at the same rate as the true sampling variability in the parameters estimates as the violation of independence increases. 
% <<simhistogrid, echo=FALSE , message=FALSE, warning=FALSE, fig.width=10, fig.height=6, out.width='1\\linewidth', fig.pos='h',fig.align='center',fig.cap="ARTIST model parameter estimates from 25,000 simulations at each $\\eta$. True parameter values from the generative model are displayed as vertical lines in each cell.", cache=TRUE>>=
% library(ggplot2)
% library(reshape2)
% etas <- seq(from=0, to=0.5, by=.1) 
% plotbetadata <- data.frame(eta = rep(as.character(etas),each=M))
% plotbetadata <- cbind(plotbetadata, rbind(resultlist[[1]][,c(1,2,7,8,13,14)],
%                                           resultlist[[2]][,c(1,2,7,8,13,14)],
%                                           resultlist[[3]][,c(1,2,7,8,13,14)],
%                                           resultlist[[4]][,c(1,2,7,8,13,14)],
%                                           resultlist[[5]][,c(1,2,7,8,13,14)],
%                                           resultlist[[6]][,c(1,2,7,8,13,14)]) )
% names(plotbetadata) <- c("eta","beta[11]","beta[12]","beta[21]","beta[22]","tau[1]","tau[2]")
% plotbetadata$eta <- paste("eta==",plotbetadata$eta,sep="")
% meltplotbetadata <- melt(plotbetadata, id=c("eta"))
% trueeffects <- unique(meltplotbetadata[,1:2])
% trueeffects$value <-  rep(c(0.046132,0.037915,0.017485, 0.008211,0,0),each=6)
% number_ticks <- function(n) {function(limits) pretty(limits, n)}
% qplot(value, geom="histogram", data=meltplotbetadata,fill=I("gray42")) + 
%   facet_grid(eta~variable, scales = "free",labeller = label_parsed) +
%   geom_vline(aes(xintercept = value), data=trueeffects, color="black")+ theme_bw()+
%   ylab("") + xlab("Estimated Parameter Values") +
%   scale_x_continuous(breaks=number_ticks(3))
% @
% \begin{table}[hbtp]
% \centering
% \begin{tabular}{ccccc} \hline
% $\eta$ & $s_{\tau_1}$ & $\hat{\text{SE}}\left[\tau_1\right]$ & $s_{\tau_2}$ & $\hat{\text{SE}}\left[\tau_2\right]$\\ 

@@

% \Sexpr{sprintf("%.1f",round(simresults[1,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[1,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[1,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[1,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[1,17],4))} \\
% \Sexpr{sprintf("%.1f",round(simresults[2,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[2,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[2,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[2,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[2,17],4))} \\
% \Sexpr{sprintf("%.1f",round(simresults[3,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[3,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[3,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[3,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[3,17],4))} \\
% \Sexpr{sprintf("%.1f",round(simresults[4,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[4,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[4,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[4,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[4,17],4))} \\
% \Sexpr{sprintf("%.1f",round(simresults[5,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[5,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[5,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[5,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[5,17],4))} \\
% \Sexpr{sprintf("%.1f",round(simresults[6,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[6,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[6,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[6,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[6,17],4))} \\
% \end{tabular}
% \caption{Comparison of the standard deviation for the treatment effect estimates, $s_{\tau_k}$, and average from standard errors for the treatment effects, $\hat{\text{SE}}\left[\tau_k\right]$ from simulations under each $\eta$.}
% \label{tab:simvarcompare}
% \end{table}
% %
% Table~\ref{tab:errorrates} displays that the Type I error rates from the Pillai's test and each of the individual t-tests for curriculum treatment effects are stable around the nominal $\alpha = 0.05$ level for all $\eta$ values. Thus the violation of independence between student responses is not detrimental to the Type I error rates in this setting, where no treatment effects exists in our generative model.  The stability in the Type I errors is linked to the finding that the variability estimates used in testing are shrinking at approximately the same rate as the true sampling variability in the parameters estimates, as discussed above.
% Figure~\ref{fig:visual1} displays that the treatment effect estimates remain unbiased and the Type I error rates for testing the significance of the treatment on each response remain at the nominal level. We may additionally be curious about the Type II error rates; failing to find effects significant that truly exist in our generative model.  Figure~\ref{fig:visual2} displays that the midterm effect estimates are unbiased for the non-zero parameters in our generative model.  It also shows that the Type II error rate in fact \textit{decreases} as the dependence between student responses increases.  This implies that power for finding a truly significant effect \textit{increased} as the violation of independence increased.  
% Our simulations, under this generative model, have demonstrated that parameter estimates remain unbiased and error rates are not detrimentally effected by increasing dependence between student responses.  Thus concerns over violating the modeling assumption of independence between student responses are alleviated. 
% \begin{table}[hbtp]
% \centering
% \begin{tabular}{cccc} \hline
% $\eta$  & Pillai's Test & t-test (CI) & t-test (HT) \\
% & $\text{H}_o : \tau_1 = \tau_2 = 0$  & $\text{H}_o : \tau_1 = 0$ &  $\text{H}_o : \tau_2 = 0$\\ 
% \Sexpr{sprintf("%.1f",round(errorsummary[1,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[1,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[1,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[1,6],4))} \\
% \Sexpr{sprintf("%.1f",round(errorsummary[2,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[2,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[2,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[2,6],4))} \\
% \Sexpr{sprintf("%.1f",round(errorsummary[3,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[3,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[3,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[3,6],4))} \\
% \Sexpr{sprintf("%.1f",round(errorsummary[4,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[4,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[4,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[4,6],4))} \\
% \Sexpr{sprintf("%.1f",round(errorsummary[5,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[5,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[5,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[5,6],4))} \\
% \Sexpr{sprintf("%.1f",round(errorsummary[6,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[6,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[6,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[6,6],4))} \\
% \end{tabular}
% \caption{Type I error rates for tests for curriculum treatment effect from 25,000 simulations at each $\eta$. All tests were conducted with a nominal significance level of $\alpha = 0.05$.}
% \label{tab:errorrates}
% \end{table}
% <<visual1, echo=FALSE , message=FALSE, warning=FALSE, fig.width=11, fig.height=3, out.width='.9\\linewidth', fig.pos='h',fig.align='center',fig.cap="Average curriculum treatment effects on each value in the simulated response pairs (left) and corresponding Type I error rates (right) from 25,000 simulations at each $\\eta$. The generative model set $\\tau_1 = \\tau_2 = 0$ and tests for significance were run using $\\alpha = 0.05$">>=
% library(ggplot2)
% library(reshape2)
% library(gridExtra)
% srm <- melt(simresults, id.var="eta", measure.var=c("trt1", "trt2"))
% levels(srm$variable) <- c("ConfMC","HypMC")
% p1 <- ggplot(aes(x=eta, y=value, group=variable), data=srm) + 
%   geom_hline(yintercept=0.00, colour="grey50") + theme_bw() + 
%   geom_point(size=4, aes(shape=variable)) + geom_line(aes(ltype=variable)) + ylim(c(-0.01,0.01)) + 
%   ylab("Estimated Treatment Effect") + xlab(expression(eta)) +
%   theme(legend.position="none")
